---
title: "Association Rule Mining"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(shinyMatrix)
library(plotly)

library(arules)
```



```{r,echo=FALSE}
sidebarPanel(
inputPanel(

  
checkboxInput("ex","Uncheck if you want to use your dataset",value = TRUE),
fileInput("file", "Upload the *.csv file with headers"),
selectInput("format","Choose format of the dataset",choices = c("matrix","database"),selected = "database"),
uiOutput("vx"),
  uiOutput("vy")

),
inputPanel(

numericInput("support","Enter minimum support ratio",value = 0.2,step = 0.01),
numericInput("confidence","Enter minimum confidenc ratio", value = 0.5,step =0.01),
numericInput("minlen","Enter minimum length", value = 2,step =1,min = 2),
numericInput("nrules","Enter how many rules", value = 1,step =1)

),
inputPanel(
  numericInput("nitems","Enter how many items", value = 1,step =1),
  downloadButton("downloadPlot", "Download ItemFrequency Plot")
),
inputPanel(
numericInput("nshelfs","Enter how many shelfs", value = 2,step =1),
downloadButton("downloadPlot2", "Download ItemShelfPlot")
)
)
mainPanel(
  h6("Enter the Item Names (on the first row margin) and transactionID(on the first column),Edit the binary values between"),
uiOutput("mat"),
numericInput("nrows","Enter number of transactions",value = 2),
numericInput("ncols","Enter number of items", value = 3),
downloadButton("downloaddata", "Download Dataset in  Matrix Format"),
br(),
br(),
tabsetPanel(type = "tab",
              tabPanel("Rules",verbatimTextOutput("AssRules")),
              tabPanel("Rules-Visuals",plotlyOutput("AssRulesV")),
              tabPanel("ItemFrequencyPlot",plotOutput("ItemFP")),
              tabPanel("Shelf-Design",plotOutput("AffV"))
               )

               
)
h6("", tags$img(src ="K.JPG", height= 400, width=400))




output$mat <- renderUI({
     matrixInput(
       "myMatrix",
       value = matrix(c(1,0,0,1,0,0),input$nrows,input$ncols),

       rows = list(names= TRUE,editableNames = TRUE),
       cols = list( names = TRUE,editableNames = TRUE),
       copy = TRUE,
       paste = TRUE,
       class = 'numeric'
     )
   })
output$AssRules <- renderPrint({
  
     
     
  
     CT =  matrix(input$myMatrix,nrow = input$nrows, ncol = input$ncols)
     row.names(CT) = row.names(input$myMatrix)
     colnames(CT) = colnames(input$myMatrix)

     
     if(input$ex == TRUE)    
     {   txn = as(CT, "transactions")
     
      

cat(sprintf("\n The details of the rules formed are as follows\n"))
rules =  arules::apriori(txn,parameter = list(support = input$support, confidence= input$confidence,minlen = input$minlen,target = "rules"))
arules::inspect(head(sort(rules,by = "lift"),n =input$nrules))

     
     }
     
     else
     {
      
       if(input$format == "matrix" && input$ex == FALSE)
       {
          file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         ds =  data.frame(data)
         CT = as.matrix(ds)
        txn = as(CT, "transactions")
        

cat(sprintf("\n The details of the rules formed are as follows\n"))
rules = arules::apriori(txn,parameter = list(support= input$support, confidence= input$confidence,minlen = input$minlen,target = "rules"))
arules::inspect(head(sort(rules,by = "lift"),n =input$nrules))

       }
       else
       {
       if(input$format == "database" && input$ex == FALSE)
       {
         file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         indexproduct= grep(input$variablex, colnames(data))
         indextrans= grep(input$variabley, colnames(data))
         splitdataset = split(data[,indexproduct],data[,indextrans])
         txn = as(splitdataset,"transactions")
         

cat(sprintf("\n The details of the rules formed are as follows\n"))
rules = arules:: apriori(txn,parameter = list(support = input$support, confidence= input$confidence,minlen = input$minlen,target = "rules"))
arules::inspect(head(sort(rules,by = "lift"),n =input$nrules))

       } 
       }
     }
     

    

     
   })

output$AssRulesV <- renderPlotly({
    CT =  matrix(input$myMatrix,nrow = input$nrows, ncol = input$ncols)
     row.names(CT) = row.names(input$myMatrix)
     colnames(CT) = colnames(input$myMatrix)

     
     if(input$ex == TRUE)    
     {   txn = as(CT, "transactions")
     
      rules = arules:: apriori(txn,parameter = list(support = input$support, confidence = input$confidence,minlen = input$minlen,target = "rules"))
arulesViz::plotly_arules(head(sort(rules,by = "lift"),n =input$nrules))
     
     }
     
     else
     {
      
       if(input$format == "matrix" && input$ex == FALSE)
       {
          file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         ds =  data.frame(data)
         CT = as.matrix(ds)
        txn = as(CT, "transactions")
        
rules = arules:: apriori(txn,parameter = list(support = input$support, confidence = input$confidence,minlen = input$minlen,target = "rules"))
arulesViz::plotly_arules(head(sort(rules,by = "lift"),n =input$nrules))
       }
       else
       {
       if(input$format == "database" && input$ex == FALSE)
       {
         file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         indexproduct= grep(input$variablex, colnames(data))
         indextrans= grep(input$variabley, colnames(data))
         splitdataset = split(data[,indexproduct],data[,indextrans])
         txn = as(splitdataset,"transactions")
         
rules = arules::apriori(txn,parameter = list(support = input$support, confidence = input$confidence,minlen = input$minlen,target = "rules"))
arulesViz::plotly_arules(head(sort(rules,by = "lift"),n =input$nrules))
       } 
       }
     }
     



  
   })

output$ItemFP <- renderPlot({
    CT =  matrix(input$myMatrix,nrow = input$nrows, ncol = input$ncols)
     row.names(CT) = row.names(input$myMatrix)
     colnames(CT) = colnames(input$myMatrix)

     
     if(input$ex == TRUE)    
     {   txn = as(CT, "transactions")
     
    
  arules::itemFrequencyPlot(txn, topN=input$nitems, type="absolute", main="Item Frequency")   
     }
     
     else
     {
      
       if(input$format == "matrix" && input$ex == FALSE)
       {
          file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         ds =  data.frame(data)
         CT = as.matrix(ds)
        txn = as(CT, "transactions")
        

arules::itemFrequencyPlot(txn, topN=input$nitems, type="absolute", main="Item Frequency") 
       }
       else
       {
       if(input$format == "database" && input$ex == FALSE)
       {
         file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         indexproduct= grep(input$variablex, colnames(data))
         indextrans= grep(input$variabley, colnames(data))
         splitdataset = split(data[,indexproduct],data[,indextrans])
         txn = as(splitdataset,"transactions")
         

arules::itemFrequencyPlot(txn, topN=input$nitems, type="absolute", main="Item Frequency") 
       } 
       }
     }
     



  
   })


output$AffV <- renderPlot({
  CT =  matrix(input$myMatrix,nrow = input$nrows, ncol = input$ncols)
     row.names(CT) = row.names(input$myMatrix)
     colnames(CT) = colnames(input$myMatrix)

     
     if(input$ex == TRUE)    
     {   txn = as(CT, "transactions")
     
      
 distance = arules::dissimilarity(txn,which = "items",method = "dice")
fit = hclust(distance,method = "ward.D")
plot(fit,main = "Dendogram of items")
rect.hclust(fit,k=input$nshelfs,border = "red")
 
     }
     
     else
     {
      
       if(input$format == "matrix" && input$ex == FALSE)
       {
          file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         ds =  data.frame(data)
         CT = as.matrix(ds)
        txn = as(CT, "transactions")
        


distance = arules::dissimilarity(txn,which = "items",method = "dice")
fit = hclust(distance,method = "ward.D")
plot(fit,main = "Dendogram of items")
rect.hclust(fit,k=input$nshelfs,border = "red")
 
       }
       else
       {
       if(input$format == "database" && input$ex == FALSE)
       {
         file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         indexproduct= grep(input$variablex, colnames(data))
         indextrans= grep(input$variabley, colnames(data))
         splitdataset = split(data[,indexproduct],data[,indextrans])
         txn = as(splitdataset,"transactions")
         

distance = arules::dissimilarity(txn,which = "items",method = "dice")
fit = hclust(distance,method = "ward.D")
plot(fit,main = "Dendogram of items")
rect.hclust(fit,k=input$nshelfs,border = "red")
       } 
       }
     }
      


   })
     
 datasetInput1 <- reactive({
  CT =  matrix(input$myMatrix,nrow = input$nrows, ncol = input$ncols)
  row.names(CT) = row.names(input$myMatrix)
  colnames(CT) = colnames(input$myMatrix)
  dataset =  data.frame(CT)
 
     
})
output$downloaddata <- downloadHandler(
    filename = function() {
      filetitle = paste("dataset")
      paste(filetitle, ".csv", sep = "")
    },
    content = function(file) {
      
      write.csv(datasetInput1(), file, row.names = FALSE)
    }
  )


output$vx <- renderUI({
   
 
     
     file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
    
     
     if(input$format == "database")
   { selectInput("variablex","Select the ItemID column",choices =   
                   colnames(data),selected = colnames(data))}
    
    
  


  })
output$vy <- renderUI({
   
 
     
     file1 = input$file
     if(is.null(file1)){return()}
     
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
    
     ds = dplyr:: select(data,-c(input$variablex))
     
     
     if(input$format == "database")
   { selectInput("variabley","Select the Transaction ID column",choices =   
                   colnames(ds),selected = colnames(ds))}
    
  

  })





output$downloadPlot<- downloadHandler(
    filename = function() {
      paste("ItemFrequencyPlot", ".png", sep = "")
    },
    content = function(file) {
      png(file)
 CT =  matrix(input$myMatrix,nrow = input$nrows, ncol = input$ncols)
     row.names(CT) = row.names(input$myMatrix)
     colnames(CT) = colnames(input$myMatrix)

     
     if(input$ex == TRUE)    
     {   txn = as(CT, "transactions")
     
      
arules::itemFrequencyPlot(txn, topN=input$nitems, type="absolute", main="Item Frequency") 
     
     }
     
     else
     {
      
       if(input$format == "matrix" && input$ex == FALSE)
       {
          file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         ds =  data.frame(data)
         CT = as.matrix(ds)
        txn = as(CT, "transactions")
        

arules::itemFrequencyPlot(txn, topN=input$nitems, type="absolute", main="Item Frequency") 
       }
       else
       {
       if(input$format == "database" && input$ex == FALSE)
       {
         file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         indexproduct= grep(input$variablex, colnames(data))
         indextrans= grep(input$variabley, colnames(data))
         splitdataset = split(data[,indexproduct],data[,indextrans])
         txn = as(splitdataset,"transactions")
         

arules::itemFrequencyPlot(txn, topN=input$nitems, type="absolute", main="Item Frequency") 
       } 
       }
     }
     



       
      dev.off()
    }
    )




output$downloadPlot2<- downloadHandler(
    filename = function() {
      paste("ItemShelfPlot", ".png", sep = "")
    },
    content = function(file) {
      png(file)
 CT =  matrix(input$myMatrix,nrow = input$nrows, ncol = input$ncols)
     row.names(CT) = row.names(input$myMatrix)
     colnames(CT) = colnames(input$myMatrix)

     
     if(input$ex == TRUE)    
     {   txn = as(CT, "transactions")
     
      
distance = arules::dissimilarity(txn,which = "items",method = "dice")
fit = hclust(distance,method = "ward.D")
plot(fit,main = "Dendogram of items")
rect.hclust(fit,k=input$nshelfs,border = "red")
     
     }
     
     else
     {
      
       if(input$format == "matrix" && input$ex == FALSE)
       {
          file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         ds =  data.frame(data)
         CT = as.matrix(ds)
        txn = as(CT, "transactions")
        

distance = arules::dissimilarity(txn,which = "items",method = "dice")
fit = hclust(distance,method = "ward.D")
plot(fit,main = "Dendogram of items")
rect.hclust(fit,k=input$nshelfs,border = "red") 
       }
       else
       {
       if(input$format == "database" && input$ex == FALSE)
       {
         file1 = input$file
     if(is.null(file1)){return()}
     data =  read.table(file =  file1$datapath,sep =",",header = TRUE)
     if(is.null(data())){return()}
         indexproduct= grep(input$variablex, colnames(data))
         indextrans= grep(input$variabley, colnames(data))
         splitdataset = split(data[,indexproduct],data[,indextrans])
         txn = as(splitdataset,"transactions")
         

distance = arules::dissimilarity(txn,which = "items",method = "dice")
fit = hclust(distance,method = "ward.D")
plot(fit,main = "Dendogram of items")
rect.hclust(fit,k=input$nshelfs,border = "red")
       } 
       }
     }
     



       
      dev.off()
    }
    )